---
title: "Field-level Completeness/Frequency"
author: "Project Team 2"
date: "2024-06-06"
output: html_document
---

#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here) 
library(geojsonR) 
library(sf) 
library(leaflet)
```

## Loading Data

```{r}
data_file_path <- here('data', 'aggregate_data.csv')
data <- read_csv(data_file_path)

geo_file_path <- here('data', 'jurisdiction_sf.geojson')
geo_data <- st_read(geo_file_path)

geo_data <- geo_data %>%  
  rename(system = sys_abbr)
```

#Fixing system N/A values
```{r}
data <- data %>%
  separate(file, into = c("press_release_system", "press_release", "year"), sep = "_") %>%
  mutate(c_system_abbr = case_when(
    is.na(c_system_abbr) ~ press_release_system,
    TRUE ~ c_system_abbr
  ))

data %>% 
  filter(is.na(c_system_abbr))
```

#DCRA required fields 
1. Decedent's name
2. Date of birth
3. Gender
4. Race 
5. Ethnicity
6. Date of death 
7. Time of death
8. Location of death
9. Law enforcement/correctional agency involved
10. Description of death (narrative and cause of death) 

```{r}
dcra_variables <- c("c_system_abbr", "c_ind_full_name", "c_ind_dob_year", "c_ind_gender", "c_ind_race", "c_ind_ethnicity", "c_ind_dod_ymd", "ind_tod", "ind_deathloc", "c_ind_fachoused", "c_ind_cod_type")
```

#DCRA required fields dataset
```{r}
#selected dataset variables based on DCRA required fields 
data <- data %>%  
  mutate(c_ind_first = 
           case_when(
             c_ind_first == "N/A" ~ NA, 
             TRUE ~ c_ind_first),
         c_ind_last = 
           case_when(
             c_ind_last == "N/A" ~ NA,
             TRUE ~ c_ind_last)
  )

dcra_dataset <- data %>% 
  mutate(c_ind_full_name = ifelse(!is.na(c_ind_first) & !is.na(c_ind_last),
                                  paste(c_ind_first, " ", c_ind_last), 
                                  NA)) %>% 
  select(all_of(dcra_variables))
```

#Total present data percentage by system and DCRA field  
```{r}
#computes the percentage of present data values for each system 
threshold <- 90 

dcra_table_system <- dcra_dataset %>% 
  group_by(c_system_abbr) %>% 
  summarize(
    present_data = sum(!is.na(across(everything()))),
    total_data = n() * (ncol(.) - 1),
    total_present_percent = round((present_data/total_data) * 100, 2)
  ) %>% 
  rename(system = c_system_abbr) %>% 
  select(system, total_present_percent) %>% 
  mutate(present = case_when(
    total_present_percent > threshold ~ "Present", 
    TRUE ~ "Missing"
  ))
dcra_table_system
```

#Total present data percentage by DCRA field 
```{r} 
#computes the percentage of present data values for each DCRA field 
dcra_table_field <- dcra_dataset %>% 
  summarize(across(everything(), ~round(mean(!is.na(.)) * 100, 2))) %>% 
  select(all_of(c(dcra_variables))) %>% 
  pivot_longer(everything(), names_to = "variables", values_to = "total_present_percent") 
dcra_table_field
```

#Total present data percentage by DCRA field grouping (demographics, cause of death, context) 
```{r} 
#computes the percentage of present data values for each DCRA field grouping 
demographic_variables <- c("c_ind_full_name", "c_ind_dob_year", "c_ind_gender", "c_ind_race", "c_ind_ethnicity") 
cod_variables <- c("c_ind_cod_type") 
context_variables <- c("c_ind_dod_ymd", "ind_tod", "ind_deathloc", "c_ind_fachoused") 

dcra_table_field_grouping <- dcra_dataset 
  
dcra_table_field_grouping 
```

#Present data percentage for each variable by system
```{r}
#computes the percentage of present data values for each variable for each system 
dcra_table_variable <- dcra_dataset %>% 
  group_by(c_system_abbr) %>% 
  summarize(across(everything(), ~ round(mean(!is.na(.)) * 100, 2))) %>% 
  select(all_of(c(dcra_variables))) %>% 
  rename(system = c_system_abbr) 
dcra_table_variable
```

#Bar Graph Data Cleaning
```{r}
threshold <- 90

dcra_table_variable_bar_cleaned <- dcra_table_variable %>% 
  mutate(c_ind_full_name = 
           case_when(
             c_ind_full_name > threshold ~ "Present", 
             TRUE ~ "Missing"),
         c_ind_dob_year = 
           case_when(
             c_ind_dob_year > threshold ~ "Present", 
             TRUE ~ "Missing"),
         c_ind_gender = 
           case_when(
             c_ind_gender > threshold ~ "Present", 
             TRUE ~ "Missing"),
         c_ind_race = 
           case_when(
             c_ind_race > threshold ~ "Present",
             TRUE ~ "Missing"),
         c_ind_ethnicity = 
           case_when(
             c_ind_ethnicity > threshold ~ "Present", 
             TRUE ~ "Missing"),
         c_ind_dod_ymd = 
           case_when(
             c_ind_dod_ymd > threshold ~ "Present", 
             TRUE ~ "Missing"),
         ind_tod = 
           case_when(
             ind_tod > threshold ~ "Present", 
             TRUE ~ "Missing"),
         ind_deathloc = 
           case_when(
             ind_deathloc > threshold ~ "Present", 
             TRUE ~ "Missing"),
         c_ind_fachoused = 
           case_when(
             c_ind_fachoused > threshold ~ "Present", 
             TRUE ~ "Missing"),
         c_ind_cod_type = 
           case_when(
             c_ind_cod_type > threshold ~ "Present", 
             TRUE ~ "Missing"),
         )
dcra_table_variable_bar_cleaned
```

#Bar Graph Dataset
```{r} 
#computes the number of systems with present/missing data for a given variable (based on 90% threshold) 
dcra_table_variable_bar <- dcra_table_variable_bar_cleaned %>% 
  pivot_longer(cols = !system, names_to = "variable", values_to = "present") %>% 
  filter(present == "Present") %>% 
  group_by(variable) %>% 
  summarize(present = n(),
            missing = 54 - present) %>% 
  ungroup() 
dcra_table_variable_bar

#computes the percentage of systems with present data for a given variable (based on 90% threshold) 
dcra_table_variable_bar_percentage <- dcra_table_variable_bar %>% 
  mutate(present_percentage = round(present/(present + missing) * 100, 2))
dcra_table_variable_bar_percentage
```

#Bar Graph
```{r}
#bar graph for the available DCRA required field data (present/missing based on 90% threshold) in press releases 
dcra_table_variable_bar %>% 
  pivot_longer(!variable, names_to = "status", values_to = "count") %>% 
  mutate(variable = factor(variable, levels = c("c_ind_full_name", "c_ind_dob_year", "c_ind_gender", "c_ind_race", "c_ind_ethnicity", "c_ind_fachoused", "ind_deathloc", "c_ind_dod_ymd", "ind_tod", "c_ind_cod_type"))) %>% 
  ggplot(aes(x = count, y = variable, fill = status)) +
  geom_bar(position = "stack", stat = "identity") + 
  geom_text(aes(label = count), position = position_stack(vjust = 0.5), color = "black", size = 2.5) + 
  labs(title = "Available DCRA Required Field Data in Press Releases", 
        caption = "DCRA required field data availability determined by a >50% \nthreshold, meaning across all press releases for a given system, \nthe data is present for at least 50% of those press releases ",
        x = "Total Number of Systems",
        y = "Required DCRA Prison Mortality Field",
        fill = "Availability for DCRA field") + 
  theme_minimal() + 
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50), limits = c(0, 55)) + 
  scale_fill_manual(values = c("#bcf6bcff", "#b8d8f0ff"), labels = c("Missing", "Present")) + 
  scale_y_discrete(labels = c("Name", "Age", "Gender", "Race", "Ethnicity", "Facility", "Location of Death", "Date of Death", "Time of Death", "Cause of Death")) + 
  theme(axis.title.y = element_text(margin = margin(r = 20))) 
```

#Join
```{r}
joined_agg_variable <- dcra_table_variable %>%  
  right_join(geo_data, join_by(system==system)) %>% 
  mutate_if(is.numeric,coalesce,0)
joined_agg_variable

joined_agg_system <- dcra_table_system %>% 
  right_join(geo_data, join_by(system==system)) %>% 
  mutate_if(is.numeric, coalesce, 0) 
joined_agg_system

joined_agg_present <- dcra_table_variable_bar_cleaned %>% 
  right_join(geo_data, join_by(system==system)) %>% 
  mutate_if(is.numeric, coalesce, 0) 
joined_agg_present
```

#Write to file/Testing 
```{r}
here()
joined_agg_variable %>%  
  st_write(here("dcra_field_completeness.geojson"), delete_dsn = T)

here()
joined_agg_system %>%  
  st_write(here("dcra_comprehensive_field_completeness.geojson"), delete_dsn = T)

here() 
joined_agg_present %>% 
  st_write(here("dcra_field_completeness_present.geojson"), delete_dsn = T) 


#write_csv(joined_agg_bar, here("dcra_field_completeness_bar.csv"))
#write_csv(dcra_table_variable_bar, here("dcra_field_completeness_count.csv")) 
```