---
title: "timelines"
author: "Project Team 2"
date: "2024-06-12"
output: html_document
---

# Load Packages

```{r}
library(tidyverse)
library(dplyr)
```

# Load Data

```{r}
data <- read_csv("/Users/emilywang/3rdcityproj/data/aggregate_data.csv")
```

# Clean data/reclassify NA systems

```{r}
data <- data |> 
  mutate(c_system_abbr = case_when(
    startsWith(file, "FL") ~ "FL", 
    startsWith(file, "AZ") ~ "AZ",
    startsWith(file, "DE") ~ "DE", 
    startsWith(file, "IA") ~ "IA",
    TRUE ~ c_system_abbr
  ))
```

# Seperate date of death into year, month, and day

```{r}
agg_data <- data |>
  drop_na(c_ind_dod_ymd) |>
  separate(c_ind_dod_ymd, into = c("dod_year", "dod_month", "dod_day"), sep = "-") |>
  group_by(dod_year) |>
  mutate(n = n()) 
```

# Create line graph with count of press releases by year

```{r}
agg_data <- agg_data |>
  arrange(dod_year)

agg_data$dod_year <- as.numeric(agg_data$dod_year)

agg_data |>
  filter(dod_year != "1950") |>
    ggplot(aes(x = dod_year, y = n, color = )) +
    geom_line(color = "#74a9cf") +
    geom_point(color = "#74a9cf") +
  labs(title = "A maximum of 1,358 press releases reported deaths that occured in 2020,
likely due to public awareness of COVID-19", x = "Year of Death", y = "Number of Press Releases Published") +
  scale_y_continuous(breaks = seq(0, 1400, 200)) +
  geom_area(fill = "#a6bddb") 
```

# Create tibble with count by year across all systems saved as releases_per_year

```{r}
releases_comp = subset(agg_data, select = c(dod_year, n))

releases_comp <- releases_comp |>
  group_by(dod_year) |>
  summarise(across(everything(), last))

write.csv(releases_comp, "../data/releases_per_year.csv")
```

# Create new variable with count by year and system

```{r}
system_count_data <- agg_data |>
  group_by(dod_year, c_system_abbr) |>
  mutate(n = n())
```

# Create final tibble with relevant counts saved as releases_per_year_per_system

```{r}
system_release_comp = subset(system_count_data, select = c(dod_year, c_system_abbr, n))

system_release_comp <- system_release_comp |>
  group_by(dod_year, c_system_abbr) |>
  summarise(across(everything(), last))

write.csv(system_release_comp, "../data/system_releases_timeline.csv")
```

# Create variable that counts number of systems that report per year

```{r}
system_release_comp <- system_release_comp |>
  group_by(dod_year) |>
  mutate(system_count = n())
```

# Create final tibble with number of systems that report per year saved as reporting_systems_per_year

```{r}
system_count_comp = subset(system_release_comp, select = c(dod_year,system_count))

system_count_comp <- system_count_comp |>
  group_by(dod_year) |>
  summarise(across(everything(), last))

write.csv(system_count_comp, "../data/reporting_systems_per_year.csv")
```

# Create tibble with the first year systems started reporting saved as system_first_release_year

```{r}
system_first_release = subset(system_count_data, select = c(dod_year, c_system_abbr))

system_first_release <- system_first_release |>
  group_by(c_system_abbr) |>
  filter(dod_year == min(dod_year)) |>
  summarise(across(everything(), last))

write.csv(system_first_release, "../data/system_first_release_year.csv")
```
