---
title: "r_webscraping_test"
output: html_document
date: "2024-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) 
library(rvest) 
library(here)
library(lubridate) 
library(chromote)
library(keyring) 
```

#Functions
```{r}
#Webscrapes all HTML tables from website URL 
scrape_table <- function(scrape_url) {
  scrape_page <- scrape_url %>% read_html() 
  
  scrape_table <- scrape_page %>% 
    html_table() %>% 
    map_dfr(~.x) %>% 
    
  return(scrape_table)
}

#Returns current date-time
date_time_current <- function() {
  time <- Sys.time() 
  return(time) 
}
```

#Florida Individual Deaths Webscraper 
```{r}
scrape_supplemental_info_fl <- function(link, session) {
  session$Page$navigate(link) 
  Sys.sleep(5) 
  
  html <- session$DOM$getDocument()$root$nodeId
  html_content <- session$DOM$getOuterHTML(nodeId = html)$outerHTML
  
  inmate_info <- read_html(html_content) %>% 
    html_nodes("table") %>% 
    .[[1]] %>%
    html_table(fill = TRUE) %>% 
    filter(!is.na(X1) & X1 != "" & !is.na(X2) & X2 != "") %>% 
    pivot_wider(names_from = X1, values_from = X2) %>% 
    mutate(ind_supplemental_link = link) 
  
  return(inmate_info)
}

scrape_individual_death_data <- function(scrape_url) {
  session <- ChromoteSession$new() 
  session$Page$navigate(scrape_url)
  Sys.sleep(5)
  
  html <- session$DOM$getDocument()$root$nodeId
  html_content <- session$DOM$getOuterHTML(nodeId = html)$outerHTML
  
  fl_scrape_tbl <- read_html(html_content) %>% 
    html_table() %>% 
    map_dfr(~.x) %>% 
    rename(ind_full_name = `Name`, 
           ind_doc_id = `DC Number`, 
           ind_dod_ymd = `Date of Death`, 
           ind_fac_housed = `Institution Name`, 
           ind_cod = `Manner of Death Determined by ME`, 
           ind_investigative_status = `Investigative Status`)
  
  fl_links <- read_html(html_content) %>% 
    html_nodes("table") %>% 
    html_nodes("a") %>% 
    map_df(~tibble(
      ind_doc_id = html_text(.), 
      ind_supplemental_link = html_attr(., "href")
    )) %>% 
    filter(ind_supplemental_link != "http://www.dc.state.fl.us/OffenderSearch/detail.aspx?Page=Detail&TypeSearch=IR&DCNumber=")
  
  fl_supplemental_table <- fl_links$ind_supplemental_link %>% 
    map_df(~scrape_supplemental_info_fl(.x, session)) %>% 
    rename(ind_doc_id = `DC Number:`, 
           ind_full_name = `Name:`, 
           ind_race = `Race:`, 
           ind_gender = `Sex:`, 
           ind_dob = `Birth Date:`, 
           ind_custody_level = `Custody:`, 
           ind_release_date = `Current Release Date:`) %>% 
    select(!`Special Note:`) %>% 
    mutate(ind_dob = as.Date(ind_dob, format = "%m/%d/%Y"))
  session$close() 
  
  fl_scrape_tbl <- fl_scrape_tbl %>% 
    left_join(fl_supplemental_table, by = "ind_doc_id") %>% 
    select(ind_supplemental_link, everything()) %>% 
    select(!ind_full_name.y) %>% 
    rename(ind_full_name = ind_full_name.x) %>% 
    mutate(ind_doc_id = 
             case_when(
               ind_doc_id == "" ~ NA, 
               TRUE ~ ind_doc_id
             ),
           ind_investigative_status = 
             case_when(
               ind_investigative_status == "" ~ NA, 
               TRUE ~ ind_investigative_status
             ),
           ind_dob = 
             case_when(
               as.character(ind_dob) == "<NA>" ~ NA, 
               TRUE ~ ind_dob
             )
           )
  fl_scrape_tbl$webscrape_date <- date_time_current()
  fl_scrape_tbl$system_abbr <- "FL" 
  
  return(fl_scrape_tbl)
}
```

```{r}
scrape_url_19_20 <- "https://www.fdc.myflorida.com/statistics-and-publications/inmate-mortality?year=2019+-+2020#mortality_graph"
fl_scrape_tbl_19_20 <- scrape_individual_death_data(scrape_url_19_20) 
fl_scrape_tbl_19_20
```

```{r}
scrape_url_20_21 <- "https://www.fdc.myflorida.com/statistics-and-publications/inmate-mortality?year=2020+-+2021#mortality_graph"
fl_scrape_tbl_20_21 <- scrape_individual_death_data(scrape_url_20_21)
fl_scrape_tbl_20_21
```

```{r}
scrape_url_21_22 <- "https://www.fdc.myflorida.com/statistics-and-publications/inmate-mortality?year=2021+-+2022#mortality_graph"
fl_scrape_tbl_21_22 <- scrape_individual_death_data(scrape_url_21_22) 
fl_scrape_tbl_21_22
```

```{r}
scrape_url_22_23 <- "https://www.fdc.myflorida.com/statistics-and-publications/inmate-mortality?year=2022+-+2023#mortality_graph"
fl_scrape_tbl_22_23 <- scrape_individual_death_data(scrape_url_22_23)
fl_scrape_tbl_22_23
```

```{r}
scrape_url_23_24 <- "https://www.fdc.myflorida.com/statistics-and-publications/inmate-mortality?year=2023+-+2024#mortality_graph"
fl_scrape_tbl_23_24 <- scrape_individual_death_data(scrape_url_23_24) 
fl_scrape_tbl_23_24
```

```{r}
scrape_url_24_25 <- "https://www.fdc.myflorida.com/statistics-and-publications/inmate-mortality?year=2024+-+2025#mortality_graph"
fl_scrape_tbl_24_25 <- scrape_individual_death_data(scrape_url_24_25) 
fl_scrape_tbl_24_25
```

```{r}
combined_fl_scrape_tbl <- bind_rows(fl_scrape_tbl_19_20, fl_scrape_tbl_20_21, fl_scrape_tbl_21_22, fl_scrape_tbl_22_23, fl_scrape_tbl_23_24, fl_scrape_tbl_24_25)
combined_fl_scrape_tbl
fl_scrape_tbl %>% write_csv(here("data/webscraping_data/florida_individual_death.csv")) 

#fl_scrape_tbl_19_20 %>% write_csv(here("data/webscraping_data/florida_individual_deaths_19_20.csv")) 
fl_scrape_tbl_20_21 %>% write_csv(here("data/webscraping_data/florida_individual_deaths_20_21.csv")) 
#fl_scrape_tbl_21_22 %>% write_csv(here("data/webscraping_data/florida_individual_deaths_21_22.csv")) 
fl_scrape_tbl_22_23 %>% write_csv(here("data/webscraping_data/florida_individual_deaths_22_23.csv")) 
#fl_scrape_tbl_23_24 %>% write_csv(here("data/webscraping_data/florida_individual_deaths_23_24.csv")) 
#fl_scrape_tbl_24_25 %>% write_csv(here("data/webscraping_data/florida_individual_deaths_24_25.csv")) 
```

#Mike Example Webscraper
```{r}
#Webscraped URL 
scrape_url = "https://cor.mt.gov/DataStatsContractsPoliciesProcedures/DataPages/DeathsInCustody"

#Read HTML content of URL 
mt_scrape_page = scrape_url %>% 
  read_html()

mt_scrape_tbl <- mt_scrape_page %>% 
  html_table() %>%  # returns all tables 
  map_dfr(~.x) %>%  # walk each table and return it
  set_names(c("ind_dod", "ind_name", "doc_id_num", "location")) %>%  # could also extract from tbl
  filter(ind_dod != "Date of Death", !is.na(ind_dod), ind_dod != "") %>%  # drop known blanks
  mutate(ind_dod_year = ind_dod %>%  mdy() %>%  year()) # create a year var for counting


mt_scrape_tbl
#mt_scrape_tbl |> write_csv("MT_death_scrape_tbl.csv") 
```

#Alabama Executions Webscraper 
```{r}
#Webscraped URL - server currently not working - NOT DONE
scrape_url <- "https://doc.alabama.gov/Executions"

al_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_doc_id", "ind_full_name", "ind_race", "ind_gender", "ind_county", "ind_offense", "ind_arrived_deathrow", "ind_dob", "ind_dod", "ind_governor")) %>% 
  mutate(ind_arrived_deathrow = 
           case_when(
             ind_arrived_deathrow == "?" ~ NA, 
             TRUE ~ ind_arrived_deathrow
            ),
         ind_dob =
           case_when(
             ind_dob == "?" ~ NA, 
             grepl("^c\\.\\s*\\d{4}", ind_dob) ~ sub("c\\.\\s*(\\d{4})", "\\1", ind_dob), 
             TRUE ~ ind_dob 
           )
  )
al_scrape_table

#al_scrape_table %>% write_csv(here("data/webscraping_data/alabama_executions.csv"))
```

#Arizona Executions Webscraper
```{r}
#Webscraped URL
scrape_url <- "https://corrections.az.gov/death-row/executions-prior-1992-execution-methods"

az_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_doc_id", "ind_full_name", "ind_ethnicity", "ind_tod", "ind_dod_ymd")) %>% 
  filter(ind_doc_id != "ADC#", ind_doc_id != "Death Penalty Inactive April 1962 - April 1992") %>% 
  mutate(ind_tod = 
           case_when(
             ind_tod == "- - - - - -" ~ NA, 
             ind_tod == "- - - - - - -" ~ NA,
             TRUE ~ ind_tod
           )
        ) %>% 
  mutate(ind_dod_ymd = as.Date(ind_dod_ymd, format = "%m-%d-%Y"))

az_scrape_table$webscrape_date <- date_time_current() 
az_scrape_table$ind_cod <- "Execution"
az_scrape_table$system_abbr <- "AZ" 
az_scrape_table

#az_scrape_table %>% write_csv(here("data/webscraping_data/arizona_executions.csv"))
```

#Arkansas Executions Webscraper
```{r}
#Webscraped URL 
scrape_url <- "https://doc.arkansas.gov/correction/inmates/executions/"

ar_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_full_name", "ind_race_gender", "ind_age", "ind_county", "ind_offense", "ind_dod_ymd")) %>% 
  filter(ind_full_name != "NAME") %>% 
  mutate(ind_age = 
           case_when(
             ind_age == "" ~ NA, 
             TRUE ~ ind_age
           )
         ) %>%
  separate(ind_race_gender, into = c("ind_race", "ind_gender"), sep = "/") %>% 
  mutate(ind_race = 
           case_when(
             ind_race == "B" ~ "Black",
             ind_race == "W" ~ "White", 
             ind_race == "Indian" ~ "Other", 
             TRUE ~ ind_race 
           ),
         ind_gender = 
           case_when(
             ind_gender == "M" ~ "Male", 
             ind_gender == "F" ~ "Female",
             TRUE ~ ind_gender
           )
         ) %>% 
  mutate(ind_dod_ymd = as.Date(ind_dod_ymd, format = "%m/%d/%Y"))
ar_scrape_table$webscrape_date <- date_time_current()
ar_scrape_table$ind_cod <- "Execution"
ar_scrape_table$system_abbr <- "AR" 
ar_scrape_table

#ar_scrape_table %>% write_csv(here("data/webscraping_data/arkansas_executions.csv"))
```

#California Executions Webscraper 
```{r}
#Webscraped URL 
scrape_url <- "https://www.cdcr.ca.gov/capital-punishment/inmates-executed-1978-to-present/"

ca_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_full_name", "ind_execution_date_received", "ind_dod_ymd", "ind_time_on_dr", "ind_age", "ind_date_state_executed")) %>% 
  mutate(ind_dod_ymd = case_when(
    !is.na(ind_date_state_executed) ~ sub("\\s*\\(.*\\)", "", ind_date_state_executed),
    TRUE ~ ind_dod_ymd
    ),
    ind_deathloc = case_when(
      !is.na(ind_date_state_executed) ~ sub(".*\\((.*)\\)", "\\1", ind_date_state_executed),
      TRUE ~ NA
    )
  ) %>% 
  select(!ind_date_state_executed) %>% 
  mutate(ind_execution_date_received = as.Date(ind_execution_date_received, format = "%m/%d/%Y"),
         ind_dod_ymd = as.Date(ind_dod_ymd, format = "%m/%d/%Y"))
ca_scrape_table

ca_links <- scrape_url %>%
  read_html() %>% 
  html_nodes("table") %>% 
  html_nodes("a") %>% 
  map_df(~tibble(
    ind_full_name = html_text(.), 
    ind_supplemental_link = html_attr(., "href")
  ))

ca_scrape_table <- ca_scrape_table %>% 
  left_join(ca_links, by = "ind_full_name") %>% 
  select(ind_supplemental_link, everything()) 
ca_scrape_table$webscrape_date <- date_time_current()
ca_scrape_table$ind_cod <- "Execution"
ca_scrape_table$system_abbr <- "CA" 
ca_scrape_table

#ca_scrape_table %>% write_csv(here("data/webscraping_data/california_executions.csv")) 
```

#Mississippi Executions Webscraper
```{r}
#Webscraped URL 
scrape_url <- "https://www.mdoc.ms.gov/general-public/death-row/death-penalty-mississippi"

ms_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_full_name", "ind_race_gender", "ind_offense", "ind_dod_ymd")) %>% 
  separate(ind_race_gender, c("ind_race", "ind_gender")) %>% 
  separate(ind_dod_ymd, into = c("month", "day", "year"), sep = "-") %>% 
  mutate(year = 
           case_when(
             year >= 30 ~ paste0("19", year),
             year < 30 ~ paste0("20", year), 
             TRUE ~ year
  )) %>% 
  mutate(ind_dod_ymd = as.Date(paste(year, month, day, sep = "-")), format = "%Y-%m-%d") %>% 
  select(!c(month, day, year, format))
ms_scrape_table$webscrape_date <- date_time_current() 
ms_scrape_table$ind_cod <- "Execution"
ms_scrape_table$system_abbr <- "MS" 
ms_scrape_table

#ms_scrape_table %>% write_csv(here("data/webscraping_data/mississippi_executions.csv"))
```

#Montana Individual Deaths Webscraper
```{r}
#Webscraped URL 
scrape_url <- "https://cor.mt.gov/DataStatsContractsPoliciesProcedures/DataPages/DeathsInCustody"

mt_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_dod_ymd", "ind_full_name", "ind_doc_id", "ind_deathloc")) %>% 
  filter(ind_dod_ymd != "Date of Death", ind_dod_ymd != "") %>% 
  separate(ind_dod_ymd, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = 
           case_when(
             year >= 30 ~ paste0("19", year), 
             year < 30 ~ paste0("20", year), 
             TRUE ~ year
           )) %>% 
  mutate(ind_dod_ymd = as.Date(paste(year, month, day, sep = "-")), format = "%Y-%m-%d") %>% 
  select(!c(month, day, year, format)) 

mt_scrape_table$webscrape_date <- date_time_current()
mt_scrape_table$system_abbr <- "MT" 
mt_scrape_table

#mt_scrape_table %>% write_csv(here("data/webscraping_data/montana_individual_death.csv"))
```

#Tennessee Executions Webscraper
```{r}
#Webscraped URL 
scrape_url <- "https://www.tn.gov/correction/statistics/executions/tennessee-executions.html"

month_names <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December") 

tn_scrape_table <- scrape_table(scrape_url) %>% 
  select(X1:X5) %>% 
  set_names(c("ind_full_name", "ind_race", "ind_offense", "ind_county", "ind_dod_ymd")) %>% 
  filter(ind_full_name != "NAME", ind_full_name != "") %>% 
  separate(ind_dod_ymd, into = c("month_day", "year"), sep = ",") %>% 
  separate(month_day, into = c("month", "day"), sep = " ") %>% 
  mutate(month = match(month, month_names),
         day = as.numeric(day), 
         year = as.numeric(year)) %>% 
  mutate(ind_dod_ymd = as.Date(paste(year, month, day, sep = "-"))) %>% 
  select(!c(month, day, year))

tn_scrape_table$webscrape_date <- date_time_current()
tn_scrape_table$ind_cod <- "Execution"
tn_scrape_table$system_abbr <- "TN"  
tn_scrape_table

#tn_scrape_table %>% write_csv(here("data/webscraping_data/tennessee_executions.csv"))
```

#Texas Executions Webscraper 
```{r}
scrape_supplemental_info <- function(link) {
  scrape_page <- read_html(link)
  
  inmate_info <- scrape_page %>% 
    html_nodes("table") %>% 
    .[[1]] %>% 
    html_table(fill = TRUE) %>% 
    filter(!is.na(X2) & X2 != "" & !is.na(X3) & X3 != "") %>% 
    pivot_wider(names_from = X2, values_from = X3) %>% 
    mutate(ind_supplemental_link = link) 
  
  return(inmate_info)
}
```

```{r}
#Webscraped URL
scrape_url <- "https://www.tdcj.texas.gov/death_row/dr_executed_offenders.html"

tx_scrape_table <- scrape_table(scrape_url) %>% 
  select(!c("Execution", "Link...2", "Link...3")) %>% 
  set_names(c("ind_last", "ind_first", "ind_doc_id", "ind_age", "ind_dod_ymd", "ind_race", "ind_county")) %>% 
  mutate(ind_doc_id = as.character(ind_doc_id)) 

tx_supplemental_inmate_info <- scrape_url %>%
  read_html() %>% 
  html_nodes("table") %>% 
  html_nodes("a") %>% 
  map_df(~tibble(
    ind_supplemental_link = html_attr(., "href")
  )) %>% 
  filter(row_number() %% 2 == 1) %>% 
  mutate(ind_supplemental_link = gsub("/death_row/", "", ind_supplemental_link), 
         ind_supplemental_link = paste0("https://www.tdcj.texas.gov/death_row/", ind_supplemental_link)) %>% 
  filter(!str_detect(ind_supplemental_link, "\\.jpg$") & !str_detect(ind_supplemental_link, "no_info_available"))

supplemental_table <- tx_supplemental_inmate_info$ind_supplemental_link %>%  
  map_df(scrape_supplemental_info) %>% 
  select(!X1) %>% 
  rename(ind_full_name = Name, 
         ind_doc_id = `TDCJ Number`, 
         ind_dob_ymd = `Date of Birth`, 
         ind_date_received = `Date Received`, 
         ind_offense_date = `Date of Offense`,
         ind_age_at_offense = `Age (at the time of Offense)`,
         ind_county = County, 
         ind_race = Race, 
         ind_gender = Gender) %>% 
  select(ind_supplemental_link, ind_doc_id, ind_dob_ymd, ind_date_received, ind_offense_date, ind_age_at_offense, ind_gender)

tx_scrape_table <- tx_scrape_table %>% 
  left_join(supplemental_table, by = "ind_doc_id") %>% 
  select(ind_supplemental_link, everything()) %>% 
  mutate(ind_dod_ymd = as.Date(ind_dod_ymd, format = "%Y/%d/%Y"),
         ind_dob_ymd = as.Date(ind_dob_ymd, format = "%m/%d/%Y"),
         ind_date_received = as.Date(ind_date_received, format = "%m/%d/%Y"), 
         ind_offense_date = as.Date(ind_offense_date, format = "%m/%d/%Y"))
tx_scrape_table$webscrape_date <- date_time_current()
tx_scrape_table$system_abbr <- "TX"
tx_scrape_table

#tx_scrape_table %>% write_csv(here("data/webscraping_data/texas_executions.csv"))
```



#Utah Individual Deaths Webscraper
```{r}
#Webscraped URL 
scrape_url <- "https://corrections.utah.gov/recently-deceased-inmates/"
ut_scrape_table <- scrape_table(scrape_url) %>% 
  set_names(c("ind_full_name", "ind_doc_id", "ind_dod_ymd", "ind_age", "ind_fachoused", "ind_cod")) %>% 
  filter(ind_full_name != "Name") %>% 
  mutate(ind_cod = 
           case_when(
             ind_cod == "N/A" ~ NA,
             TRUE ~ ind_cod
           )
         ) %>% 
  mutate(ind_dod_ymd = as.Date(ind_dod_ymd, format = "%m/%d/%Y"))

ut_scrape_table$webscrape_date <- date_time_current()
ut_scrape_table$system_abbr <- "UT" 
ut_scrape_table 

#ut_scrape_table %>% write_csv(here("data/webscraping_data/utah_individual_death.csv"))
```

